---
title: "Comparing NPDR, STIR and Relief-F"
output: html_document
---

### Load packages:

```{r echo=FALSE}
library(privateEC)
library(broom)
library(tidyverse)
# library(devtools)
# install_github("insilico/npdr") # npdr install
library(npdr)

rm(list = ls())
set.seed(1618)

cbbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
                "#0072B2", "#D55E00", "#CC79A7", "#c5679b", "#be548f")
```


### Simulate case-control interaction effect data:

```{r}
#####

##### simulate case-control interaction effect data 
n.samples <- 300     # 100 samples in train/holdout/test
n.variables <- 1000   # 100 features
label <- "qtrait"   # tells simulator to do quantitative trait and adds this colname
type <- "mainEffect"
bias <- 0.6          # moderate effect size
pct.signals <- 0.1   # pct functional features
verbose <- FALSE
qtrait.3sets <- createSimulation(num.samples = n.samples,
                              num.variables = n.variables,
                              pct.signals = pct.signals,
                              label = label,
                              bias = bias,
                              pct.train = 1/3,
                              pct.holdout = 1/3,
                              pct.validation = 1/3,
                              sim.type = type,
                              save.file = NULL,
                              verbose = verbose)
  # combine train and holdout into 200 samples x 100 attributes
  # ignore validation set for now; qtrait = quantitative outcome
qtrait.data <- rbind(qtrait.3sets$train,qtrait.3sets$holdout)
n.samples.qtrait <- dim(qtrait.data)[1]
pheno.qtrait <- as.factor(qtrait.data[,label])
functional.qtrait <- qtrait.3sets$signal.names # functional attributes
```

### Run npdr:

```{r}
#####
system.time(
  npdr.qtrait.results <- 
    npdr("qtrait", qtrait.data, regression.type="lm", 
            attr.diff.type = "numeric-abs", nbd.method = "multisurf", 
            nbd.metric = "manhattan", msurf.sd.frac = .5, 
            padj.method = "bonferroni", verbose = T) %>%
    mutate(npdr.log10 = -log10(pval.att))
)

# get functional attribute detection stats
npdr.qtrait.positives <- npdr.qtrait.results %>%
  filter(pval.adj < .05) %>%
  pull(att)
npdr.qtrait.detect.stats <- detectionStats(functional.qtrait, npdr.qtrait.positives)
cat(npdr.qtrait.detect.stats$report)

```

### Univariate regression:

```{r}
system.time(
gene_mdd <- uniReg(
  outcome = label, 
  dataset = qtrait.data, 
  regression.type = 'lm') %>%
  data.frame() %>%
  rownames_to_column('att') %>%
  mutate(functional = grepl('sim', att))
)
```

### Comparing univariate regression and NPDR:

```{r}
n.pos <- gene_mdd %>%
  filter(p.adj < 0.05) %>%
  nrow()
true.pos <- gene_mdd %>%
  filter((p.adj < 0.05) & functional) %>%
  nrow()
(uni.recall <- true.pos/(n.variables*pct.signals))
(uni.precis <- true.pos/n.pos)

test.df <- npdr.qtrait.results %>%
  mutate(functional = att %in% functional.qtrait) 
true.pos <- test.df %>%
  filter((pval.adj < 0.05) & functional) %>%
  nrow()
n.pos <- test.df %>%
  filter((pval.adj < 0.05)) %>%
  nrow()
(npdr.recall <- true.pos/(n.variables*pct.signals))
(npdr.precis <- true.pos/n.pos)

pr_uni <- PRROC::pr.curve(scores.class0 = gene_mdd %>%
                        filter(functional == T) %>%
                        pull(beta.Z.att) %>% abs(),
                      scores.class1 = gene_mdd %>%
                        filter(functional == F) %>%
                        pull(beta.Z.att) %>% abs(),
                      curve = T)

pr_npdr <- PRROC::pr.curve(scores.class0 = test.df %>%
                        filter(functional == T) %>%
                        pull(beta.Z.att),
                      scores.class1 = test.df %>%
                        filter(functional == F) %>%
                        pull(beta.Z.att),
                      curve = T)
pr_uni$auc.integral
pr_npdr$auc.integral
```


### Compute Relief-F score with the `CORElearn` package
(fixed $k$ with theoretical surf value)

```{r}
core.learn.qtrait <- CORElearn::attrEval(label, data = qtrait.data,
                                      estimator = "RReliefFequalK",
                                      costMatrix = NULL,
                                      outputNumericSplits=FALSE,
                                      kNearestEqual = knnSURF(n.samples.qtrait,.5))
corelearn.df <- data.frame(att=names(core.learn.qtrait), rrelief=core.learn.qtrait)
last.imp.att <- which(npdr.qtrait.results$pval.adj>.05)[1]-1
npdr.beta.cutoff.qtrait <- npdr.qtrait.results[last.imp.att, 'beta.Z.att']

test.df <- merge(corelearn.df, npdr.qtrait.results) %>%
  mutate(functional = att %in% functional.qtrait)
r2_qtrait <- round(cor(test.df$rrelief, test.df$beta.Z.att), digits = 3)
```


### Draw the Precision-Recall curve:

```{r}
pr_relief <- PRROC::pr.curve(scores.class0 = test.df %>%
                        filter(functional == T) %>%
                        pull(rrelief),
                      scores.class1 = test.df %>%
                        filter(functional == F) %>%
                        pull(rrelief),
                      curve = T)

x <- as.data.frame(pr_npdr$curve)
colnames(x) <- paste0(c('Recall', 'Precision', 'w'))
y <- as.data.frame(pr_relief$curve)
colnames(y) <- paste0(c('Recall', 'Precision', 'w'))
auprc_npdr_qtrait <- round(pr_npdr$auc.integral, 3)
auprc_relief_qtrait <- round(pr_relief$auc.integral, 3)
prr.dat <- rbind(x, y) %>%
  mutate(type = rep(c('NPDR', 'Relief-F'), each = nrow(x)))
```



What about auROC?

```{r}
ss_npdr <- pROC::roc(test.df$functional, test.df$beta.Z.att)
ss_relief <- pROC::roc(test.df$functional, test.df$rrelief)
auroc_npdr_qtrait <- round(ss_npdr$auc, 3)
auroc_relief_qtrait <- round( ss_relief$auc, 3)

ss_df_qtrait <- data.frame(Sensitivity = c(ss_npdr$sensitivities, ss_relief$sensitivities),
                    Specificity = c(ss_npdr$specificities, ss_relief$specificities)) %>%
  mutate(type = rep(c('NPDR', 'Relief'), each = length(ss_npdr$sensitivities)))
```


How much better is NPDR? Is this just by chance? 100 replications result show the same thing (npdr_100_quant.Rmd).

```{r}
# ss_df_qtrait <- ss_df
test_df_qtrait <- test.df
prr_dat_qtrait <- prr.dat

save(ss_df_qtrait, test_df_qtrait, prr_dat_qtrait,
     auprc_npdr_qtrait, auprc_relief_qtrait, r2_qtrait,
     npdr.beta.cutoff.qtrait, ss_df_qtrait,
     auroc_npdr_qtrait, auroc_relief_qtrait,
     file = 'results/df_qtrait.Rdata')

```



